---
title: "ENIGMA-22q Normative Centile Generation"
author: "Benjamin Jung, PhD"
format:
  pdf:
    documentclass: scrreprt
    toc: true
    number-sections: true
execute:
  echo: true
  warning: false
  message: false
editor: visual
---

```{r}
#| label: setup-environment
#| output: false
rm(list=ls())
library(tidyverse)
library(glue)
library(freesurfer)
library(ggseg)
# Loads Calc.Novel(): applies a pre-fit LBCC GAMLSS model to new ("novel") data
# out-of-sample
source('../Lifespan/940.calc-novel-wo-subset-function.r')
```

## Load Pre-Fit LBCC Models

```{r}
#| label: compile-lbcc-feature-lists
#| output: false
# Read in global model features
features.glob <- basename(Sys.glob("../../../../braincharts/LBCC/global/RDS/*Transformed"))
features.glob <- str_replace_all(features.glob,"globalv1-","")
features.glob <- str_replace_all(features.glob,"Transformed","")

# Read in SA model features
features.sa <- basename(Sys.glob("../../../../braincharts/LBCC/test-SA-site-merge-v2/RDS/*.SA.*"))
features.sa <- str_replace_all(features.sa,"site240920-","")
features.sa <- str_replace_all(features.sa,"Transformed","")

# Read in CT model features
features.ct <- basename(Sys.glob("../../../../braincharts/LBCC/test-CT-site-merge-v5/RDS/*.CT.*"))
features.ct <- str_replace_all(features.ct,"site240920-","")
features.ct <- str_replace_all(features.ct,"Transformed","")

# Read in GM model features
features.gm <- basename(Sys.glob("../../../../braincharts/LBCC/test-GM-site-merge-v3/RDS/*.GM.*"))
features.gm <- str_replace_all(features.gm,"site240920-","")
features.gm <- str_replace_all(features.gm,"Transformed","")

# Read in subcort GM model features
features.subc <- basename(dirname(Sys.glob("../../../../braincharts/LBCC/test-SUBC-site-merge-v2/RDS/*SUBC.*/FIT.EXTRACT.rds")))
features.subc <- str_replace_all(features.subc,"site240920-","")
features.subc <- str_replace_all(features.subc,"Transformed","")

features <- c(features.subc, features.glob,features.gm, features.ct, features.sa)

```

```{r}
#| label: exclude-invalid-subc-features
#| output: false
features <- features[!(features %in% c("SUBC.Left.choroid.plexus", "SUBC.Right.choroid.plexus","SUBC.CSF","SUBC.5th.Ventricle","SUBC.Optic.Chiasm"))]
```

## Freesurfer Volume Processing (R)

### Compile Output

```{r}
#| label: set-freesurfer-dir
#| output: false
# Specify directory w/ FS outputs
fs_dir <- "../../../../Data/misc_research/22q-ENIGMA/subjects"
```

```{r}
#| label: process-aparc-cortical-stats
#| output: false
# Read in LH cortical structures
lh.aparc.stats <- read_atlas_files(fs_dir, "lh.aparc.stats")
lh.aparc.stats$prefix <- "lh"
# Read in RH cortical structures
rh.aparc.stats <- read_atlas_files(fs_dir, "rh.aparc.stats")
rh.aparc.stats$prefix <- "rh"

# Filter unused cortical features
exclude_features <- c("NumVert", "ThickStd","GausCurv","FoldInd","CurvInd","MeanCurv")

# Concatenate cortical features
aparc.stats <- bind_rows(lh.aparc.stats,rh.aparc.stats)
aparc.stats <- aparc.stats[,!(colnames(aparc.stats) %in% exclude_features)]

# Wide to long
aparc.stats <- gather(aparc.stats, measure, value, SurfArea:ThickAvg, factor_key=TRUE)

# Preliminary relabeling
aparc.stats <- aparc.stats %>%
                    mutate(merge_label = glue("{prefix}.{measure}_{label}")) %>%
                    mutate(merge_label = str_replace_all(merge_label,
                                                         c('SurfArea_' = "SA.",
                                                            'ThickAvg_' = "CT.",
                                                            'GrayVol_' = "GM.",
                                                            'lh_' = "",
                                                            'rh_' = "")))


aparc.stats <- aparc.stats[,!(colnames(aparc.stats) %in% c("prefix"))]
```

```{r}
#| label: process-aseg-subcortical-stats
#| output: false
# Read in SUBCortical structures
aseg.stats <- read_atlas_files(fs_dir, "aseg.stats")

# Filter SUBCortical structures
aseg.stats <- aseg.stats %>%
                filter(!grepl("-vessel",label)) %>%
                filter(!grepl("hypointensities",label))

exclude_features <- c("NVoxels", "Index","SegId","FoldInd","normMean","normStdDev","normMin","normRange","normMax")
aseg.stats <- aseg.stats[,!(colnames(aseg.stats) %in% exclude_features)]
# Rename columns for consistency
colnames(aseg.stats) <- c("subject","Vol","label")

# Convert wide to long
aseg.stats <- gather(aseg.stats, measure, value, Vol, factor_key=TRUE)

# Preliminary relabeling
aseg.stats <- aseg.stats %>%
                    mutate(merge_label = glue("SUBC.{label}")) %>%
                    mutate(merge_label = str_replace_all(merge_label,
                                                        c('-' = '.',
                                                          "\\.inf\\." = ".inferior.",
                                                          "\\.lat\\." = ".lateral.",
                                                          "\\.vent$" = ".ventricle")))
```

```{r}
#| label: compute-global-stats-per-subject
#| output: false
#Loop through each subject to read global measures
fs_sub_paths <- Sys.glob(glue("{fs_dir}/*"))
for (fs_sub_path in fs_sub_paths) {
    sub = basename(fs_sub_path)

    # Load in global features (aseg)
    file <- glue("{fs_sub_path}/aseg.stats")
    exclude_measures <- c("brainsegnotventsurf","mask",
                          "surfaceholes","maskvol-to-etiv",
                          "brainsegvol-to-etiv")
    if (file.exists(file)) {
      global.stats.ind <- read_aseg_stats(file)$measures %>%
                      add_column(subject = sub, .before = "measure") %>%
                      filter(!(measure %in% exclude_measures))
    }
    # Load in global features (lh)
    file <- glue("{fs_sub_path}/lh.aparc.stats")
    if (file.exists(file)) {
      lh_stats <- read_aseg_stats(file)
      lh_sa <- sum(lh_stats$structures$SurfArea)
      global.stats.ind <- bind_rows(global.stats.ind,lh_stats$measures %>%
                      add_column(subject = sub, .before = "measure") %>%
                      filter(!(measure_long %in% c("numvert"))) %>%
                      mutate(measure_long = str_replace_all(measure_long,
                                              c("whitesurfarea" = "lhwhitesurfacearea",
                                                "meanthickness" = "lhmeanthickness"))))
    }


    # Load in global features (rh)
    file <- glue("{fs_sub_path}/rh.aparc.stats")
    if (file.exists(file)) {
      rh_stats <- read_aseg_stats(file)
      rh_sa <- sum(rh_stats$structures$SurfArea)
      global.stats.ind <- bind_rows(global.stats.ind,rh_stats$measures %>%
                      add_column(subject = sub, .before = "measure") %>%
                      filter(!(measure_long %in% c("numvert"))) %>%
                      mutate(measure_long = str_replace_all(measure_long,
                                              c("whitesurfarea" = "rhwhitesurfacearea",
                                                "meanthickness" = "rhmeanthickness"))))
    }
    global.stats.ind <- global.stats.ind %>%
                          mutate(value = as.numeric(value)) %>%
                          add_row(subject = sub,
                                  measure = "SA",
                                  measure_long = "totalsurfarea",
                                  meaning = "total surface area",
                                  value = lh_sa + rh_sa,
                                  units = "mm^2")

    # Store global stats
    if (fs_sub_path == fs_sub_paths[1]) {
        global.stats <- global.stats.ind
    } else {
        global.stats <- bind_rows(global.stats,global.stats.ind)
    }
}

global.stats <- tibble(global.stats)

# Remove extraneous columns
global.stats <- global.stats[!(colnames(global.stats) %in% c("measure","meaning","units"))]
#Relabel for consistency
colnames(global.stats) <- c("subject","label","Vol")
# Wide to long
global.stats <- gather(global.stats, measure, value, Vol, factor_key=TRUE)

# Preliminary relabeling
global.stats <- global.stats %>%
                    mutate(merge_label = glue("GLOB.{tolower(label)}")) %>%
                    mutate(merge_label = str_replace_all(merge_label,
                                                      c("\\.lh" = ".left.",
                                                        "\\.rh" = ".right.")))
```

```{r}
#| label: compute-euler-number
#| output: false
euler.stats <- global.stats %>%
                filter(grepl("surfacehole",label)) %>%
                mutate(euler = 2-2*as.numeric(value)) %>%
                group_by(subject) %>%
                summarize(euler = mean(euler))
```

```{r}
#| label: combine-regional-and-global-measures
#| output: false
#Concatenate regional measures
region.stats <- bind_rows(aparc.stats,aseg.stats)
#Concatenate global measures
region.stats$value <- as.double(region.stats$value)
global.stats$value <- as.double(global.stats$value)
mri.measures <- bind_rows(region.stats,global.stats)
mri.measures <- mri.measures[,!(colnames(mri.measures) %in% c("label","measure"))]

mri.measures <- mri.measures[!duplicated(mri.measures),]
```

### Reformat Structure Names

```{r}
#| label: rename-global-structure-labels
#| output: false
mri.measures <- mri.measures %>%
                  mutate(merge_label = replace(merge_label,
                                               merge_label == "GLOB.etiv",
                                               "eTIV"),
                         merge_label = replace(merge_label,
                                               merge_label == "GLOB.totalsurfarea",
                                               "SA"))
```

```{r}
#| label: compute-composite-regions
#| output: false
# Convert from long to wide
mri.measures.wide <- spread(mri.measures, merge_label, value)

# Compute composite regions
# ventricles.all
mri.measures.wide$Ventricles.all <- mri.measures.wide$SUBC.Left.Lateral.Ventricle +
                                    mri.measures.wide$SUBC.Left.Inf.Lat.Vent +
                                    mri.measures.wide$SUBC.Right.Lateral.Ventricle +
                                    mri.measures.wide$SUBC.Right.Inf.Lat.Vent +
                                    mri.measures.wide$SUBC.3rd.Ventricle +
                                    mri.measures.wide$SUBC.4th.Ventricle

# ventricles.lateral
mri.measures.wide$Ventricles.lateral <- mri.measures.wide$SUBC.Left.Lateral.Ventricle +
                                        mri.measures.wide$SUBC.Right.Lateral.Ventricle

# Thalamus
mri.measures.wide$Thalamus <- mri.measures.wide$SUBC.Left.Thalamus.Proper +
                              mri.measures.wide$SUBC.Right.Thalamus.Proper

# cerebellum
mri.measures.wide$Cerebellum.GM <- mri.measures.wide$SUBC.Left.Cerebellum.Cortex +
                                   mri.measures.wide$SUBC.Right.Cerebellum.Cortex

mri.measures.wide$Cerebellum.WM <- mri.measures.wide$SUBC.Left.Cerebellum.White.Matter +
                                   mri.measures.wide$SUBC.Right.Cerebellum.White.Matter

mri.measures.wide$Cerebellum.all <- mri.measures.wide$Cerebellum.GM +
                                    mri.measures.wide$Cerebellum.WM

# CT
mri.measures.wide$CT <- mri.measures.wide %>%
                            select(GLOB.left.meanthickness,GLOB.right.meanthickness) %>%
                            rowMeans()
```

## Meta and Demographic Data Processing

### Import manifest file

```{r}
#| label: load-manifest-file
#| output: false
# Specify path of pre-existing manifest
manifest_file <- "../data/ENIGMA/ENIGMA_22q_data.csv"
 
# Load pre-existing manifest
manifest_df <- read_csv(manifest_file, show_col_types = FALSE)
 
qc_cols <- c("general white matter", "general grey matter", "general csf", "cerebellum",
             "brainstem", "thalamus", "putamen+pallidum", "hippocampus+amygdala")
```

### Get resolution and QC info

```{r}
#| label: compute-qc-metrics
#| output: false

manifest_df$qc_path <- NA
manifest_df[qc_cols] <- NA
 
for (i in 1:nrow(manifest_df)) {
  # Get scan path and qc for each subject
  subj_id <- manifest_df$SubjID[i]
  scan_matches <- Sys.glob(file.path(fs_dir, subj_id, "mri", "orig", "*.mgz"))
 
  if (length(scan_matches) == 1) {
    qc_matches <- Sys.glob(file.path(fs_dir, subj_id, "T1_ss_qc.csv"))
 
    if (length(qc_matches) == 1) {
      manifest_df$qc_path[i] <- qc_matches[1]
      qc_df <- read_csv(qc_matches[1], n_max = 1, show_col_types = FALSE)
      manifest_df[i, qc_cols] <- qc_df[1, qc_cols]
    }
  }
}
 
# Flag QC status: TRUE if all metrics > 0.65, FALSE if any metric is <= 0.65
# or missing, NA if every metric is missing
manifest_df <- manifest_df %>%
  rowwise() %>%
  mutate(QC = {vals <- c_across(all_of(qc_cols))
                if (all(is.na(vals))) {
                  NA
                } else {
                  all(replace(vals, is.na(vals), -Inf) > 0.65)
                }
              }) %>%
  ungroup()
```

### create participant, session, site, dx, age_days, mean_qc columns

```{r}
#| label: create-demographic-columns
#| output: false
# Rename columns, derive age/session, and recode diagnosis
manifest_df <- manifest_df %>%
  rename(participant = SubjID,
          site        = Site,
          dx          = DxName,
          sex         = SexName) %>%
  mutate(age_days = Age * 365.25 + 280) %>%
  group_by(participant) %>%
  mutate(session = paste0("ses-", dense_rank(age_days))) %>%
  ungroup() %>%
  mutate(
    magnetic_strength = NA,
    mean_qc = rowMeans(across(all_of(qc_cols)), na.rm = TRUE),
    dx = if_else(dx == "HC", "CN", dx)
  )
```

### Merge manifest with morphometrics

```{r}
#| label: merge-manifest-with-morphometrics
#| output: false
# Merge in imaging measures and drop redundant columns
# (assumes `mri.measures.wide` is already an R data frame in your environment,
# replacing the reticulate `r["mri.measures.wide"]` reference)
df <- manifest_df %>%
  left_join(mri.measures.wide, by = c("participant" = "subject")) %>%
  select(-any_of(c("Frank_MergeID", "Dx", "Sex", "subject")))
```

### Save df

```{r}
#| label: save-enigma-volumes-csv
#| output: false
# Save output
dir.create("../data/ENIGMA/")
write_csv(df, "../data/ENIGMA/ENIGMA_volumes.csv")
```

## Predict Centiles (R)

### Load Data

```{r}
#| label: load-and-rename-global-columns
#| output: false
df = read.csv("../data/ENIGMA/ENIGMA_volumes.csv")
# Rename global structures
glob.measures <- grepl("GLOB.",colnames(df))
colnames(df)[glob.measures] <- colnames(df)[glob.measures] %>%
                                          str_replace_all("cerebellum","Cerebellum") %>%
                                          str_replace_all("ventricles","Ventricles") %>%
                                          str_replace_all("thalamus","Thalamus") %>%
                                          str_replace_all("brainstem","Brainstem") %>%
                                          str_replace_all("gm","GM") %>%
                                          str_replace_all("wm","WM") %>%
                                          str_replace_all("GLOB.","")
```

```{r}
#| label: prepare-covariates-and-save-raw
#| output: false
# Add hard-coded variables
df$sex = factor(df$sex,levels=c('Female','Male'),labels=c('Female',"Male"))
df$fs_version = 'FS53'
df$AgeTransformed=log(df$age_days)
df$study <- toupper(df$Mega_Scanner_Name)
df$study[is.na(df$study)] <- toupper(df$site[is.na(df$study)])
df$site <- df$study

df <- df %>%
            select(-c(general.white.matter, general.grey.matter, general.csf,
                      cerebellum, brainstem, thalamus, putamen.pallidum,
                      hippocampus.amygdala)) %>%
            filter(dx %in% c("CN","22q11DS")) %>%
            mutate(dx = factor(dx, levels = c("CN","22q11DS")),
                   study_site = study)

```

### Predict Centiles

```{r}
#| label: apply-lbcc-models-to-regions
#| output: false
# For each region, load the corresponding pre-fit LBCC GAMLSS model (feature
# set determined by which features.* list the region belongs to) and apply it
# out-of-sample via Calc.Novel to obtain centile scores (*.q.wre) and
# normalised deviation scores (*.normalised) for the ENIGMA-22q subjects.
event_list <- list()
MODELS <- list()

exclude_list <- c("SUBC.Left.Cerebellum.White.Matter",
                  "SUBC.Right.Cerebellum.White.Matter",
                  "Brainstem","SUBC.Left.vessel","SUBC.Right.vessel",
                  "SUBC.non.WM.hypointensities","SUBC.WM.hypointensities",
                  "SUBC.CC_Anterior", "SUBC.CC_Central",
                  "SUBC.CC_Mid_Anterior","SUBC.CC_Mid_Posterior",
                  "SUBC.CC_Posterior","Thalamus","SUBC.5th_Ventricle",
                  "Ventricles.lateral","Cerebellum.GM","Cerebellum.WM",
                  "Ventricles.lateral")

retain_cols <- c("participant", "age_days", "sex",'dx.original','site',
                 "Psychotic_Disorder", "Standard_Handedness", "Within_Group_Scanner",
                 "Mega_Scanner", "Deletion_Type", "IQ", "IQ_Method", "Psychotic_Disorder_Name",
                 "Current_Typical_Antipsychotic", "Current_Li", "Current_AntiConvulsant",
                 "Current_AntiDep", "Current_Psychostimulant", "Current_Other_Med")
#Subset to event
NewData <- df[df$session == "ses-1" & !(df$study %in% c("UTRECHT", "TORONTO2")),]
region_list <- list()
norm_list <- list()
for(region in features) {
  if (region %in% exclude_list) {
    next
  } else if (region %in% features.glob) {
    rds_dir <- "../../../../braincharts/LBCC/global/RDS/globalv1-"
    fit_file <- paste0(rds_dir, region, "Transformed/FIT.EXTRACT.rds")
    derived_file <- paste0(rds_dir, region, "Transformed/DERIVED.without.boot.rds")
    NewData$fs_version <- "FS5.3"
  } else if (region %in% features.subc) {
    rds_dir <- "../../../../braincharts/LBCC/test-SUBC-site-merge-v2/RDS/site240920-"
    fit_file <- paste0(rds_dir, region, "Transformed/FIT.EXTRACT.rds")
    derived_file <- paste0(rds_dir, region, "Transformed/DERIVED.without.boot.v2.rds")
    NewData$fs_version <- "FS5.3"
  } else if (region %in% features.gm) {
    rds_dir <- "../../../../braincharts/LBCC/test-GM-site-merge-v3/RDS/site240920-"
    fit_file <- paste0(rds_dir, region, "Transformed/FIT.EXTRACT.rds")
    derived_file <- paste0(rds_dir, region, "Transformed/DERIVED.without.boot.v2.rds")
    NewData$fs_version <- "FS5.3"
  } else if (region %in% features.ct) {
    rds_dir <- "../../../../braincharts/LBCC/test-CT-site-merge-v5/RDS/site240920-"
    fit_file <- paste0(rds_dir, region, "Transformed/FIT.EXTRACT.rds")
    derived_file <- paste0(rds_dir, region, "Transformed/DERIVED.without.boot.v2.rds")
    NewData$fs_version <- "FS5.3"
  } else if (region %in% features.sa) {
    rds_dir <- "../../../../braincharts/LBCC/test-SA-site-merge-v2/RDS/site240920-"
    fit_file <- paste0(rds_dir, region, "Transformed/FIT.EXTRACT.rds")
    derived_file <- paste0(rds_dir, region, "Transformed/DERIVED.without.boot.v2.rds")
    NewData$fs_version <- "FS5.3"
  } else {
    next
  }
  if(file.exists(fit_file)) {
    if(nrow(NewData) > 1) {
      FitObj <- readRDS(fit_file)
      # Global models are currently not compatible :(
      if (is.null(attr(FitObj$param,"transformations"))) {
        next
      }
      RESULT <- Calc.Novel(NewData, Fit = FitObj, Apply = TRUE, NumberCores = 1)
      region_list[[region]] <- RESULT[["data"]][c(retain_cols, paste0(region, "Transformed.q.wre"))]
      norm_list[[region]] <- RESULT[["data"]][c(retain_cols, paste0(region, "Transformed.normalised"))]

      #Extract curves
      if (file.exists(derived_file)) {
        DATA <- readRDS(derived_file)
      } else {
        DATA <- readRDS(str_replace_all(derived_file,".without.boot.v2",""))
      }
      MODELS[[region]] <- DATA$POP.CURVE.PRED
      MODELS[[region]]$name <- region

    }
  }
}

df <- region_list %>% reduce(left_join, by = retain_cols)
norm_df <- norm_list %>% reduce(left_join, by = retain_cols)

ALL.MODELS = do.call(bind_rows,MODELS)

```

Next, we'll use the braincharts webapp to extract the global centiles

```{r}
#| label: load-global-centiles-from-webapp
#| output: false
# Global metrics (GMV, sGMV, WMV, Ventricles, meanCT, totalSA) use the LBCC
# global models, which are incompatible with direct Calc.Novel application
# (see skip above), so their centiles were generated separately via the
# braincharts webapp and are merged in here.
glob_features <- list()
glob_features[["GMV"]]  <- read_csv("../data/ENIGMA/brainchart_results/GMV_centiles_2025-01-17.csv", show_col_types = F) %>%
                              select(participant, sex, AgeTransformed,
                                     GMVTransformed.q.wre, GMVTransformed.normalised,
                                     PRED.l025.pop, PRED.m500.pop, PRED.u975.pop) %>%
                              mutate(name = "GMV")

glob_features[["sGMV"]]  <- read_csv("../data/ENIGMA/brainchart_results/sGMV_centiles_2025-01-17.csv", show_col_types = F) %>%
                              select(participant, sex, AgeTransformed,
                                     sGMVTransformed.q.wre, sGMVTransformed.normalised,
                                     PRED.l025.pop, PRED.m500.pop, PRED.u975.pop) %>%
                              mutate(name = "sGMV")

glob_features[["WMV"]]  <- read_csv("../data/ENIGMA/brainchart_results/WMV_centiles_2025-01-17.csv", show_col_types = F) %>%
                              select(participant, sex, AgeTransformed,
                                     WMVTransformed.q.wre, WMVTransformed.normalised,
                                     PRED.l025.pop, PRED.m500.pop, PRED.u975.pop) %>%
                              mutate(name = "WMV")

glob_features[["Ventricles"]]  <- read_csv("../data/ENIGMA/brainchart_results/Ventricles_centiles_2025-01-17.csv",
                                         show_col_types = F) %>%
                              select(participant, sex, AgeTransformed,
                                     VentriclesTransformed.q.wre, VentriclesTransformed.normalised,
                                     PRED.l025.pop, PRED.m500.pop, PRED.u975.pop) %>%
                              mutate(name = "Ventricles")

glob_features[["meanCT"]]  <- read_csv("../data/ENIGMA/brainchart_results/CT_centiles_2025-01-17.csv", show_col_types = F) %>%
                              select(participant, sex, AgeTransformed,
                                     meanCTTransformed.q.wre = meanCT2Transformed.q.wre,
                                     meanCTTransformed.normalised = meanCT2Transformed.normalised,
                                     PRED.l025.pop, PRED.m500.pop, PRED.u975.pop) %>%
                              mutate(name = "meanCT")


glob_features[["totalSA"]]  <- read_csv("../data/ENIGMA/brainchart_results/SA_centiles_2025-02-07.csv", show_col_types = F) %>%
                              select(participant, sex, AgeTransformed,
                                     totalSATransformed.q.wre = totalSA2Transformed.q.wre,
                                     totalSATransformed.normalised = totalSA2Transformed.normalised,
                                     PRED.l025.pop, PRED.m500.pop, PRED.u975.pop) %>%
                              mutate(name = "totalSA")
```

```{r}
#| label: merge-global-centiles
#| output: false
for (i in names(glob_features)) {
    # Add pre-calculated centiles for GMV, WMV, and sGMV
    df <- left_join(df, glob_features[[i]] %>%
                                    select(-contains("PRED."),
                                           -contains(".normalised"),
                                           -c(name, AgeTransformed)))

    # Add normalized values
    norm_df <- left_join(norm_df, glob_features[[i]] %>%
                                    select(-contains("PRED."),
                                           -contains(".q.wre"),
                                           -c(name, AgeTransformed)))

    # Add curves
    ALL.MODELS <- bind_rows(ALL.MODELS,
                            glob_features[[i]] %>%
                              select(AgeTransformed, sex, name,
                                     PRED.l025.pop, PRED.m500.pop, PRED.u975.pop))
}

df <- df %>% select(-c(Ventricles.allTransformed.q.wre))
norm_df <- norm_df %>% select(-c(Ventricles.allTransformed.normalised))
ALL.MODELS <- ALL.MODELS %>% filter(name != "Ventricles.all")
```

## Inspect Centile Distributions

```{r}
#| label: plot-global-centile-boxplots
#| fig-cap: "Centile scores for global brain measures (GMV, WMV, sGMV, ventricles, eTIV) by diagnostic group in the ENIGMA-22q cohort."
tmp_df <- df %>%
            select(participant, dx.original, age_days,
                   GMVTransformed.q.wre, WMVTransformed.q.wre,
                   sGMVTransformed.q.wre, VentriclesTransformed.q.wre,
                   eTIVTransformed.q.wre) %>%
            pivot_longer(-c(participant,dx.original, age_days),
                         names_to = "measure",values_to = "centile") %>%
            mutate(measure = str_replace_all(measure,"Transformed.q.wre",""))

# Plot global centiles by dx
ggplot(tmp_df, aes(x = dx.original, y = centile)) +
    geom_hline(yintercept = 0.5, linetype = "dashed") +
    geom_boxplot() + facet_wrap(~ measure, ncol =3)
```

```{r, fig.height = 16, fig.width = 12}
#| label: plot-sa-centile-boxplots
#| fig-cap: "Regional cortical surface area centile scores by diagnostic group in the ENIGMA-22q cohort."
tmp_df <- df %>%
            select(participant, dx.original, age_days, contains(".SA.")) %>%
            pivot_longer(-c(participant,dx.original, age_days),
                         names_to = "measure",values_to = "centile") %>%
            mutate(measure = str_replace_all(measure,"Transformed.q.wre",""))

ggplot(tmp_df, aes(x = dx.original, y = centile)) +
    geom_hline(yintercept = 0.5, linetype = "dashed") +
    geom_boxplot() + facet_wrap(~ measure, ncol =5)
```

```{r, fig.height = 16, fig.width = 12}
#| label: plot-ct-centile-boxplots
#| fig-cap: "Regional cortical thickness centile scores by diagnostic group in the ENIGMA-22q cohort."
tmp_df <- df %>%
            select(participant, dx.original, age_days, contains(".CT.")) %>%
            pivot_longer(-c(participant,dx.original, age_days),
                         names_to = "measure",values_to = "centile") %>%
            mutate(measure = str_replace_all(measure,"Transformed.q.wre",""))

ggplot(tmp_df, aes(x = dx.original, y = centile)) +
    geom_hline(yintercept = 0.5, linetype = "dashed") +
    geom_boxplot() + facet_wrap(~ measure, ncol =5)
```

```{r, fig.height = 16, fig.width = 12}
#| label: plot-gm-centile-boxplots
#| fig-cap: "Regional gray matter volume centile scores by diagnostic group in the ENIGMA-22q cohort."
tmp_df <- df %>%
            select(participant, dx.original, age_days, contains(".GM.")) %>%
            pivot_longer(-c(participant,dx.original, age_days),
                         names_to = "measure",values_to = "centile") %>%
            mutate(measure = str_replace_all(measure,"Transformed.q.wre",""))

ggplot(tmp_df, aes(x = dx.original, y = centile)) +
    geom_hline(yintercept = 0.5, linetype = "dashed") +
    geom_boxplot() + facet_wrap(~ measure, ncol =5)
```

## Final Processing

```{r}
#| label: rename-phenotype-columns
#| output: false
df <- df %>%
        rename(dx = dx.original) %>%
        mutate(Source = "ENIGMA",
               session = 1,
               Copy_Num = case_match(dx,
                                          "22q11DS"  ~ "Del",
                                          "CN"       ~ "CN")) %>%
        rename(dx_psychosis = Psychotic_Disorder_Name) %>%
        mutate(dx_psychosis = case_match(dx_psychosis,
                                         "" ~ NA,
                                         "22q11DS-Psy"  ~ 0,
                                         "22q11Dup-Psy" ~ 0,
                                         "NonPsy" ~ 0,
                                         "22q11DS+Psy" ~ 1))


colnames(df) <- colnames(df) %>%
                      str_replace_all("h\\.","h_") %>%
                      str_replace_all("Right\\.","Right_") %>%
                      str_replace_all("Left\\.","Left_") %>%
                      str_replace_all("lh.CT.","CT.lh_") %>%
                      str_replace_all("rh.CT.","CT.rh_") %>%
                      str_replace_all("lh.SA.","SA.lh_") %>%
                      str_replace_all("rh.SA.","SA.rh_") %>%
                      str_replace_all("lh.GM.","GM.lh_") %>%
                      str_replace_all("rh.GM.","GM.rh_") %>%
                      str_replace_all("Ventral.DC","VentralDC") %>%
                      str_replace_all("GLOB.Ventricles", "Ventricles") %>%
                      str_replace_all("SUBC","SUBC") %>%
                      str_replace_all("\\.Cortex","_Cortex") %>%
                      str_replace_all("\\.Vent","_Vent") %>%
                      str_replace_all("\\.Lat","_Lat") %>%
                      str_replace_all("\\.Proper","_Proper") %>%
                      str_replace_all("3rd","x3rd") %>%
                      str_replace_all("4th","x4th")

colnames(norm_df) <- colnames(norm_df) %>%
                      str_replace_all("h\\.","h_") %>%
                      str_replace_all("Right\\.","Right_") %>%
                      str_replace_all("Left\\.","Left_") %>%
                      str_replace_all("lh.CT.","CT.lh_") %>%
                      str_replace_all("rh.CT.","CT.rh_") %>%
                      str_replace_all("lh.SA.","SA.lh_") %>%
                      str_replace_all("rh.SA.","SA.rh_") %>%
                      str_replace_all("lh.GM.","GM.lh_") %>%
                      str_replace_all("rh.GM.","GM.rh_") %>%
                      str_replace_all("Ventral.DC","VentralDC") %>%
                      str_replace_all("GLOB.Ventricles", "Ventricles") %>%
                      str_replace_all("SUBC","SUBC") %>%
                      str_replace_all("\\.Cortex","_Cortex") %>%
                      str_replace_all("\\.Vent","_Vent") %>%
                      str_replace_all("\\.Lat","_Lat") %>%
                      str_replace_all("\\.Proper","_Proper") %>%
                      str_replace_all("3rd","x3rd") %>%
                      str_replace_all("4th","x4th")

# Add Euler Number
df <- df %>% left_join(euler.stats, by = c("participant" = "subject"))
```

## Combine Volumes and Centiles into Single File

```{r}
#| label: combine-volumes-and-centiles
#| output: false
final_df <- left_join(df,norm_df) %>%
                relocate(Source:dx.original, .after = dx) %>%
                select(-c(dx.original, Copy_Num))
```

## Save ENIGMA Analysis Dataset

```{r}
#| label: save-output-csvs
#| output: false
write.csv(final_df, '../data/ENIGMA/ENIGMA_data.csv', row.names = F)
```

## Session Info

```{r}
#| label: session-info
sessionInfo()
```
