---
title: "Case-Control Effect Sizes and Sensitivity Analyses (Figure 1)"
author: "Benjamin Jung, PhD"
format:
  pdf:
    documentclass: scrreprt
    toc: true
    number-sections: true
execute:
  echo: true
  warning: false
  message: false
editor: visual
---

```{r setup}
#| echo: false
#| output: false
#| label: setup
source("util_scripts/utils.R")
library(foreach)
library(doParallel)
library(doRNG)
```

```{r}
#| label: load-analysis-data
#| output: false
load('../data/analysis_prep_mpr.RData')
df_chop_mpr <- df %>% 
                filter(Source == "CHOP")

load('../data/analysis_prep_combined.RData')
df_chop <- df %>% 
              filter(Source == "CHOP")
df_enigma <- df %>% 
              filter(Source == "ENIGMA")

# Revision: 07/21/2026 - Add race/ethnicity
df_race <- read_csv("../data/CHOP/race_ethnicity_07212026.csv") %>% distinct()
df_chop <- left_join(df_chop %>% select(-race), df_race, by = c("participant" = "subject_id"))

# Add squared-age for sensitivity analysis
df_chop <- df_chop %>% mutate(age_days_sq = age_days**2)
df_chop_mpr <- df_chop_mpr %>% mutate(age_days_sq = age_days**2)
df_enigma <- df_enigma %>% mutate(age_days_sq = age_days**2)
```

## Primary Analysis: Case-Control Deviation Scores (Figure 1)

### CHOP

```{r}
#| output: false
#| label: fit-models-chop-zscore
cov_str <- " + age_days + sex + euler"
rand_str <- ""

mpr_only_cols <- features[grepl("CT",features)]
time_start <- Sys.time()
res <- fit_group_comparison_models(df_chop, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)
time_end <- Sys.time()

print(time_end - time_start)

res_chop_z <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-chop-zscore
print_top_significant_results(res_chop_z)
summarize_results_by_cortex(res_chop_z, caption = "Case-Control Results Grouped by Cortex in CHOP")
```

#### Generate Formatted Report Table

```{r}
#| label: define-fmt-table-function
#| output: false
fmt_table <- function(res) {
  res_fmt <- res %>% 
                  select(label, contains("Cohens_d"),statistic, std.error,
                         starts_with("p."), Case, Control) %>%
                  mutate(label = case_match(label,
                                            "Cerebellum.all" ~ "Cerebellum",
                                            "SUBC.Left_Accumbens.area" ~ "SUBC.Left_Accumbens",
                                            "SUBC.Right_Accumbens.area" ~ "SUBC.Right_Accumbens",
                                            .default = label)) %>%
                  separate_wider_delim(label, delim = ".", names = c("IDP Group", "IDP"),
                                       too_few = "align_end") %>%
                  mutate(`IDP Group` = ifelse(is.na(`IDP Group`), "Global", `IDP Group`),
                         `IDP Group` = factor(`IDP Group`, levels = c("Global","SUBC","GM","SA","CT")),
                         across(contains("Cohens"), ~ round(.x, 2)),
                         across(starts_with("p."), ~ signif(.x, digits = 3)),
                         statistic = round(statistic, 2),
                         std.error = round(std.error, 2),
                         `Cohen's d (95% CI)` = glue("[{Cohens_d_ci_lo} to {Cohens_d_ci_hi}]")) %>%
                  select(-c(Cohens_d_ci_lo,Cohens_d_ci_hi)) %>%
                  relocate(`Cohen's d (95% CI)`, .after = Cohens_d) %>%
                  rename(`Cohen's d` = Cohens_d,
                         Statistic = statistic,
                         `Std Error` = std.error,
                         p = p.value,
                         `p (FDR)` = p.fdr,
                         `N case` = Case,
                         `N control` = Control) %>%
                  arrange(`IDP Group`, IDP)
  return(res_fmt)
}
```

```{r}
#| label: save-table-chop-zscore
#| output: false
table_fmt <- fmt_table(res_chop_z)
dir.create("../tables/formatted/", recursive = T)
write_csv(table_fmt, "../tables/formatted/table_s03.csv")
```

### ENIGMA

```{r}
#| output: false
#| label: fit-models-enigma-zscore
cov_str <- " + age_days + sex + euler"
rand_str <- ""

mpr_only_cols <- c()

res <- fit_group_comparison_models(df_enigma, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)

res_enigma_z <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-enigma-zscore
print_top_significant_results(res_enigma_z)
summarize_results_by_cortex(res_enigma_z, caption = "Case-Control Results Grouped by Cortex in ENIGMA")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-enigma-zscore
#| output: false
table_fmt <- fmt_table(res_enigma_z)
write_csv(table_fmt, "../tables/formatted/table_s04.csv")
```

### Permutation Test

#### CHOP

```{r}
#| label: define-permutation-parameters-zscore
#| output: false
# Define cluster parameters
n_cores <- 72
n_perm = 1000

# Define perm features
perm_features <- features[startsWith(features,"CT.") | startsWith(features,"GM.") | startsWith(features,"SA.")]
```

```{r}
#| label: run-permutation-test-chop-zscore
#| output: false
perms <- freedman_lane_perm(df_chop, perm_features, n_perm = n_perm, n_core = n_cores, suf = "Transformed.q.zscore")

true_brain_map <- res_chop_z %>% select(label, Cohens_d)
comp_brain_map <- res_enigma_z %>% select(label, Cohens_d)

res <- freedman_lane_corr_test(true_brain_map, perms, comp_brain_map)
```

### Generate Plots

```{r}
#| label: define-plot-params-zscore
#| output: false
fill_col <- "Cohens_d"
ci_lo_col <- "Cohens_d_ci_lo"
ci_hi_col <- "Cohens_d_ci_hi"
sig_col  <- "sig"
colorbar_title <- "Cohen's d"
max_lim <- ceiling(max(abs(res_chop_z$Cohens_d), abs(res_enigma_z$Cohens_d))*10)/10
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)

max_lim_main <- ceiling(max(abs(res_chop_z$Cohens_d))*10)/10
plot_lims_main <- c(-1*max_lim_main,max_lim_main)
```

```{r}
#| output: false
#| label: build-brain-maps-zscore

p_chop_z <- assemble_main_figure_layout(res_chop_z, df_chop, fill_col, color_scheme, plot_lims,
                 plot_type = "zscore", colorbar_title = colorbar_title,sig_col = sig_col,
                 formatted = T)

p_enigma_z <- assemble_main_figure_layout(res_enigma_z, df_enigma, fill_col, color_scheme, plot_lims,
                 plot_type = "zscore", colorbar_title = colorbar_title,sig_col = sig_col,
                 formatted = T)
```

```{r}
#| label: build-main-figure-and-corr-plot-zscore
#| output: false

p_chop_z_main <- assemble_main_figure_layout(res_chop_z, df_chop, fill_col, color_scheme, plot_lims_main,
                 plot_type = "zscore", colorbar_title = colorbar_title,sig_col = sig_col,
                 formatted = T)

tmp <- inner_join(res_chop_z %>% select(label, Cohens_d),
           res_enigma_z %>% select(label, Cohens_d),
           by = "label") %>%
      filter(grepl("GM\\.|SA\\.|CT\\.",label)) %>%
      separate_wider_delim(label, delim = ".", names = c("measure","label")) %>%
      left_join(res) %>%
      mutate(measure = factor(measure, levels = c("GM","SA","CT"),
                              labels = c("Gray Matter Volume", "Surface Area", "Cortical Thickness")),
             corr = as.character(round(corr,2)),
             pval = as.character(round(pval,4)),
             label_text = paste0(
                              "r == ", corr,
                              "~','~p[perm] == ", pval),
             label_text = str_replace_all(label_text, "== 0$", "< 0.001"),
             label_text = ifelse(label == "lh_bankssts",label_text, ""))

p_corr <- ggplot(tmp, aes(x = Cohens_d.x, y = Cohens_d.y)) + 
              geom_point() + # geom_smooth(method = "lm", se = F) +
              geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
              facet_wrap(~ measure, nrow = 1) +
              theme_pubr() +
              # coord_fixed(ratio = 1) +
              xlab("Cohen's d (CHOP)") +
              ylab("Cohen's d (ENIGMA)") +
              geom_text(aes(label = label_text, x=-0.7, y=0.95), parse = T, size = annotation_pt(7)) # +
```

```{r, fig.width = 6.69, fig.height = 7.5}
#| echo: false
#| label: chop-effect
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the CHOP cohort."

ggarrange(p_chop_z_main, p_corr + figure_theme, ncol = 1, heights = c(5, 2.5),
          labels = c("","D"), font.label = list(size = 10, face = "bold", color = "black"))
dir.create("../figures_final/png/", recursive = T, showWarnings = F)
dir.create("../figures_final/pdf/", recursive = T, showWarnings = F)
ggsave("../figures_final/png/figure_1_case_control_chop.png", bg = 'white', dpi = 600)
ggsave("../figures_final/pdf/figure_1_case_control_chop.pdf", bg = 'white')
```

```{r, fig.width = 6.69, fig.height = 7.5}
#| echo: false
#| label: enigma-effect
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the ENIGMA Cohort."
ggarrange(p_enigma_z, p_corr, ncol = 1, heights = c(1, 0.5),
          labels = c("","D)"))
ggsave("../figures_final/png/figure_s3_case_control_enigma.png", bg = 'white')
ggsave("../figures_final/pdf/figure_s3_case_control_enigma.pdf", bg = 'white')
```

#### Combine Plots

```{r, fig.width = 6.69, fig.height = 12.1}
#| label: combine-plots-zscore
#| fig-cap: "Combined CHOP and ENIGMA case-control effect sizes (Cohen's d) for normative deviation z-scores, with between-cohort permutation-test correlations."
t1 <- ggdraw() + 
  draw_label("CHOP", x = 0.01, hjust = 0, fontface = "bold")
t2 <- ggdraw() + 
  draw_label("ENGIMA", x = 0.01, hjust = 0, fontface = "bold")

gm_text <- glue('Gray Matter Volume: r = {round(res[res$measure == "GM","corr"][[1]],2)}, p = {round(res[res$measure == "GM","pval"][[1]],4)}')
sa_text <- glue('Surface Area: r = {round(res[res$measure == "SA","corr"][[1]],2)}, p = {round(res[res$measure == "SA","pval"][[1]],4)}')
ct_text <- glue('Cortical Thickness: r = {round(res[res$measure == "CT","corr"][[1]],2)}, p = {round(res[res$measure == "CT","pval"][[1]],4)}')

caption <- ggdraw() + 
              draw_label(glue('{gm_text} | {sa_text} | {ct_text}'))

p <- ggarrange(t1, t2, p_chop_z, p_enigma_z, nrow = 4, heights = c(0.08,1,0.08,1))
ggarrange(p, caption, heights = c(11.6,0.5), ncol = 1)
dir.create("../figures_final/")
```

### Save Effect-Size Tables for Downstream Scripts

```{r}
#| label: save-results-zscore
#| output: false
write_csv(res_chop_z, "../tables/chop_cohens.csv")
write_csv(res_enigma_z, "../tables/enigma_cohens.csv")
```

## Sensitivity Analysis: Adjusting for Intracranial Volume (Figure S6)

*Sensitivity analysis: covarying for intracranial-volume (eTIV) deviation score.*

### CHOP

```{r}
#| output: false
#| label: fit-models-chop-etiv
cov_str <- " + age_days + sex + euler + eTIVTransformed.q.zscore"
rand_str <- ""

mpr_only_cols <- features[grepl("CT",features)]

res <- fit_group_comparison_models(df_chop, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)

res_chop_etiv <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-chop-etiv
print_top_significant_results(res_chop_etiv)
summarize_results_by_cortex(res_chop_etiv, caption = "Case-Control Results Grouped by Cortex in CHOP")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-chop-etiv
#| output: false
table_fmt <- fmt_table(res_chop_etiv)
write_csv(table_fmt, "../tables/formatted/table_s05.csv")
```

### ENIGMA

```{r}
#| output: false
#| label: fit-models-enigma-etiv

mpr_only_cols <- c()

res <- fit_group_comparison_models(df_enigma, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)

res_enigma_etiv <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-enigma-etiv
print_top_significant_results(res_enigma_etiv)
summarize_results_by_cortex(res_enigma_etiv, caption = "Case-Control Results Grouped by Cortex in ENIGMA")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-enigma-etiv
#| output: false
table_fmt <- fmt_table(res_enigma_etiv)
write_csv(table_fmt, "../tables/formatted/table_s06.csv")
```

### Permutation Test

#### Generate CHOP Permutations

```{r}
#| label: run-permutation-test-etiv
#| output: false
# ADDITIONAL COVS
perms <- freedman_lane_perm(df_chop, perm_features, n_perm = n_perm, 
                            n_core = n_cores, suf = "Transformed.q.zscore",
                            covs = "age_days + sex + euler + eTIVTransformed.q.zscore")

true_brain_map <- res_chop_etiv %>% select(label, Cohens_d)
comp_brain_map <- res_enigma_etiv %>% select(label, Cohens_d)

res <- freedman_lane_corr_test(true_brain_map, perms, comp_brain_map)
```

### Generate Plots

```{r}
#| label: define-plot-params-etiv
#| output: false
fill_col <- "Cohens_d"
ci_lo_col <- "Cohens_d_ci_lo"
ci_hi_col <- "Cohens_d_ci_hi"
sig_col  <- "sig"
colorbar_title <- "Cohen's d"
max_lim <- max(abs(res_chop_etiv$Cohens_d), abs(res_enigma_etiv$Cohens_d))
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)
```

```{r}
#| output: false
#| label: build-brain-maps-etiv

p_chop_etiv <- assemble_main_figure_layout(res_chop_etiv, df_chop, fill_col, color_scheme, plot_lims,
                 plot_type = "zscore", colorbar_title = colorbar_title,sig_col = sig_col)

p_enigma_etiv <- assemble_main_figure_layout(res_enigma_etiv, df_enigma, fill_col, 
                               color_scheme, plot_lims,plot_type = "zscore",
                               colorbar_title = colorbar_title,sig_col = sig_col)
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: chop-effect-etiv
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the CHOP cohort."
p_chop_etiv
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: enigma-effect-etiv
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the ENIGMA Cohort."
p_enigma_etiv
```

#### Combine Plots

```{r, fig.width = 6.69, fig.height = 11}
#| label: combine-plots-etiv
#| fig-cap: "Combined CHOP and ENIGMA case-control effect sizes (Cohen's d) covarying for intracranial volume (eTIV), with between-cohort permutation-test correlations."
t1 <- ggdraw() + 
  draw_label("CHOP", x = 0.01, hjust = 0, fontface = "bold")
t2 <- ggdraw() + 
  draw_label("ENGIMA", x = 0.01, hjust = 0, fontface = "bold")

gm_text <- glue('Gray Matter Volume: r = {round(res[res$measure == "GM","corr"][[1]],2)}, p = {round(res[res$measure == "GM","pval"][[1]],4)}')
sa_text <- glue('Surface Area: r = {round(res[res$measure == "SA","corr"][[1]],2)}, p = {round(res[res$measure == "SA","pval"][[1]],4)}')
ct_text <- glue('Cortical Thickness: r = {round(res[res$measure == "CT","corr"][[1]],2)}, p = {round(res[res$measure == "CT","pval"][[1]],4)}')

caption <- ggdraw() + 
              draw_label(glue('{gm_text} | {sa_text} | {ct_text}'), size = base_font) + figure_theme

p <- ggarrange(t1, p_chop_etiv, t2, p_enigma_etiv, nrow = 4, heights = c(0.08,1,0.08,1))
ggarrange(p, caption, heights = c(11.6,0.5), ncol = 1)
dir.create("../figures_final/")
ggsave("../figures_final/png/figure_s6_sensitivity_icv.png", bg = 'white')
ggsave("../figures_final/pdf/figure_s6_sensitivity_icv.pdf", bg = 'white')
```

### Describe Changes

```{r}
#| label: describe-changes-etiv
# --- swap this to summarize a different sensitivity result -------------------
res_test <- res_chop_etiv    # tibble to summarize
res_ref  <- res_chop_z       # reference (primary) result to compare against
# -----------------------------------------------------------------------------

summarize_model_changes(res_test, res_ref)
```

## 

## Sensitivity Analysis: High-Resolution MPRAGE Scans Only (Figure S17)

### CHOP

```{r}
#| output: false
#| label: fit-models-chop-mprage
cov_str <- " + age_days + sex + euler"
rand_str <- ""

mpr_only_cols <- c()

res <- fit_group_comparison_models(df_chop_mpr, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)

res_chop_mpr <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-chop-mprage
print_top_significant_results(res_chop_mpr)
summarize_results_by_cortex(res_chop_mpr, caption = "Case-Control Results Grouped by Cortex in CHOP")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-chop-mprage
#| output: false
table_fmt <- fmt_table(res_chop_mpr)
write_csv(table_fmt, "../tables/formatted/table_s6.csv")
```

### Permutation Test

#### Generate CHOP Permutations

```{r}
#| label: run-permutation-test-mprage
#| output: false
perms <- freedman_lane_perm(df_chop_mpr, perm_features, n_perm = n_perm, n_core = n_cores, suf = "Transformed.q.zscore")

true_brain_map <- res_chop_mpr %>% select(label, Cohens_d)
comp_brain_map <- res_enigma_z %>% select(label, Cohens_d)

res <- freedman_lane_corr_test(true_brain_map, perms, comp_brain_map)
```

### Generate Plots

```{r}
#| label: define-plot-params-mprage
#| output: false
fill_col <- "Cohens_d"
ci_lo_col <- "Cohens_d_ci_lo"
ci_hi_col <- "Cohens_d_ci_hi"
sig_col  <- "sig"
colorbar_title <- "Cohen's d"
max_lim <- max(abs(res_chop_mpr$Cohens_d), abs(res_enigma_z$Cohens_d))
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)
```

```{r}
#| output: false
#| label: build-brain-map-mprage

p_chop_mpr <- assemble_main_figure_layout(res_chop_mpr, df_chop_mpr, fill_col, color_scheme, plot_lims,
                 plot_type = "zscore", colorbar_title = colorbar_title,sig_col = sig_col)
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: chop-effect-mprage
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the CHOP cohort (MPRAGE only)."
p_chop_mpr
```

#### Combine Plots

```{r, fig.width = 6.69, fig.height = 5}
#| label: combine-plots-mprage
#| echo: false
#| fig-cap: "Case-control effect sizes (Cohen's d) in the CHOP cohort restricted to MPRAGE acquisitions."
p_chop_mpr
dir.create("../figures_final/")
ggsave("../figures_final/png/figure_s17_sensitivity_mprage.png", bg = 'white')
ggsave("../figures_final/pdf/figure_s17_sensitivity_mprage.pdf", bg = 'white')
```

### Describe Changes

```{r}
#| label: describe-changes-mprage
# --- swap this to summarize a different sensitivity result -------------------
res_test <- res_chop_mpr    # tibble to summarize
res_ref  <- res_chop_z       # reference (primary) result to compare against
# -----------------------------------------------------------------------------

summarize_model_changes(res_test, res_ref)
```

## Sensitivity Analysis: Brain Centiles Instead of Deviation Scores (Figure S18)

*Sensitivity analysis: raw centiles in place of normative deviation (z-)scores.*

### CHOP

```{r}
#| output: false
#| label: fit-models-chop-centile
cov_str <- " + age_days + sex + euler"
rand_str <- ""

mpr_only_cols <- features[grepl("CT",features)]

res <- fit_group_comparison_models(df_chop, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "centile",
                 mpr_only_cols = mpr_only_cols)

res_chop <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-chop-centile
print_top_significant_results(res_chop)
summarize_results_by_cortex(res_chop, caption = "Case-Control Results Grouped by Cortex in CHOP")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-chop-centile
#| output: false
table_fmt <- fmt_table(res_chop)
write_csv(table_fmt, "../tables/formatted/table_s17.csv")
```

### ENIGMA

```{r}
#| output: false
#| label: fit-models-enigma-centile
cov_str <- " + age_days + sex + euler"
rand_str <- ""

mpr_only_cols <- c()

res <- fit_group_comparison_models(df_enigma, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "centile",
                 mpr_only_cols = mpr_only_cols)

res_enigma <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-enigma-centile
print_top_significant_results(res_enigma)
summarize_results_by_cortex(res_enigma, caption = "Case-Control Results Grouped by Cortex in ENIGMA")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-enigma-centile
#| output: false
table_fmt <- fmt_table(res_enigma)
write_csv(table_fmt, "../tables/formatted/table_s18.csv")
```

### Permutation Test

#### Generate CHOP Permutations

```{r}
#| label: run-permutation-test-centile
#| output: false
perms <- freedman_lane_perm(df_chop, perm_features, n_perm = n_perm, n_core = n_cores, suf = "Transformed.q.wre")

true_brain_map <- res_chop %>% select(label, Cohens_d)
comp_brain_map <- res_enigma %>% select(label, Cohens_d)

res <- freedman_lane_corr_test(true_brain_map, perms, comp_brain_map)
```

### Generate Plots

```{r}
#| label: define-plot-params-centile
#| output: false
fill_col <- "Cohens_d"
ci_lo_col <- "Cohens_d_ci_lo"
ci_hi_col <- "Cohens_d_ci_hi"
sig_col  <- "sig"
colorbar_title <- "Cohen's d "
max_lim <- max(abs(res_chop$Cohens_d), abs(res_enigma$Cohens_d))
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)
```

```{r}
#| output: false
#| label: build-brain-maps-centile

p_chop <- assemble_main_figure_layout(res_chop, df_chop, fill_col, color_scheme, plot_lims,
                 plot_type = "centile", colorbar_title = colorbar_title,sig_col = sig_col)

p_enigma <- assemble_main_figure_layout(res_enigma, df_enigma, fill_col, color_scheme, plot_lims,
                 plot_type = "centile", colorbar_title = colorbar_title,sig_col = sig_col)
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: chop-effect-centile
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the CHOP cohort."
p_chop
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: enigma-effect-centile
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the ENIGMA Cohort."
p_enigma
```

#### Combine Plots

```{r, fig.width = 6.69, fig.height = 12.1}
#| label: combine-plots-centile
#| fig-cap: "Combined CHOP and ENIGMA case-control effect sizes (Cohen's d) using raw centile scores, with between-cohort permutation-test correlations."
t1 <- ggdraw() + 
  draw_label("CHOP", x = 0.01, hjust = 0, fontface = "bold")
t2 <- ggdraw() + 
  draw_label("ENGIMA", x = 0.01, hjust = 0, fontface = "bold")

gm_text <- glue('Gray Matter Volume: r = {round(res[res$measure == "GM","corr"][[1]],2)}, p = {round(res[res$measure == "GM","pval"][[1]],4)}')
sa_text <- glue('Surface Area: r = {round(res[res$measure == "SA","corr"][[1]],2)}, p = {round(res[res$measure == "SA","pval"][[1]],4)}')
ct_text <- glue('Cortical Thickness: r = {round(res[res$measure == "CT","corr"][[1]],2)}, p = {round(res[res$measure == "CT","pval"][[1]],4)}')

caption <- ggdraw() + 
              draw_label(glue('{gm_text} | {sa_text} | {ct_text}'), size = base_font) + figure_theme

p <- ggarrange(t1, p_chop, t2, p_enigma, nrow = 4, heights = c(0.08,1,0.08,1))
ggarrange(p, caption, heights = c(11.6,0.5), ncol = 1)
dir.create("../figures_final/")
ggsave("../figures_final/png/figure_s18_sensitivity_centiles.png", bg = 'white')
ggsave("../figures_final/pdf/figure_s18_sensitivity_centiles.pdf", bg = 'white')
```

### Describe Changes

```{r}
#| label: describe-changes-centile
# --- swap this to summarize a different sensitivity result -------------------
res_test <- res_chop    # tibble to summarize
res_ref  <- res_chop_z       # reference (primary) result to compare against
# -----------------------------------------------------------------------------

# summarize_model_changes() is defined in utils.R
summarize_model_changes(res_test, res_ref)
```

## Sensitivity Analysis: Non-Benchmarked Brain Features (Figure S19)

*Sensitivity analysis: raw (non-benchmarked) volume/surface-area/thickness features in place of centiles.*

```{r}
#| label: define-feature-groups-idp
#| output: false
features.sa <- c(features[startsWith(features,"SA.")],"totalSA")
features.ct <- c(features[startsWith(features,"CT.")],"meanCT")
features.gm <- features[!(features %in% c(features.sa, features.ct))]
```

### CHOP

```{r}
#| output: false
#| label: fit-models-chop-idp
rand_str <- ""

cov.gm <- " + age_days + age_days_sq  + sex + euler"
res.gm <- fit_group_comparison_models(df_chop, cov.gm, features.gm, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "volume")

cov.ct <- " + age_days_sq + sex + euler_mpr"
res.ct <- fit_group_comparison_models(df_chop, cov.ct, features.ct, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "volume")

cov.sa <- " + age_days + sex + euler"
res.sa <- fit_group_comparison_models(df_chop, cov.sa, features.sa, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "volume")

res <- bind_rows(res.gm, res.ct, res.sa)


res_chop_vol <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-chop-idp
print_top_significant_results(res_chop_vol)
summarize_results_by_cortex(res_chop_vol, caption = "Case-Control Results Grouped by Cortex in CHOP")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-chop-idp
#| output: false

res_chop_vol <- res_chop_vol %>% 
                  arrange(match(label, res_chop_z$label))
table_fmt <- fmt_table(res_chop_vol)
write_csv(table_fmt, "../tables/formatted/table_s19.csv")
```

### ENIGMA

```{r}
#| output: false
#| label: fit-models-enigma-idp
rand_str <- ""

cov.gm <- " + age_days + age_days_sq  + sex + euler"
res.gm <- fit_group_comparison_models(df_enigma, cov.gm, features.gm, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "volume")

cov.ct <- " + age_days_sq + sex + euler"
res.ct <- fit_group_comparison_models(df_enigma, cov.ct, features.ct, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "volume")

cov.sa <- " + age_days + sex + euler"
res.sa <- fit_group_comparison_models(df_enigma, cov.sa, features.sa, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "volume")

res <- bind_rows(res.gm, res.ct, res.sa)

res_enigma_vol <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-enigma-idp
print_top_significant_results(res_enigma_vol)
summarize_results_by_cortex(res_enigma_vol, caption = "Case-Control Results Grouped by Cortex in ENIGMA")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-enigma-idp
#| output: false

res_enigma_vol <- res_enigma_vol %>% 
                  arrange(match(label, res_enigma_z$label))
table_fmt <- fmt_table(res_enigma_vol)
write_csv(table_fmt, "../tables/formatted/table_s20.csv")
```

### Permutation Test

#### Generate CHOP Permutations

```{r}
#| label: run-permutation-test-idp
#| output: false
# THIS ONE NEEDS ADDITIONAL COVARIATES
cov.gm <- "age_days + age_days_sq  + sex + euler"
cov.ct <- "age_days_sq + sex + euler"
cov.sa <- "age_days + sex + euler"

feat.gm <- perm_features[grepl("GM\\.", perm_features)]
feat.sa <- perm_features[grepl("SA\\.", perm_features)]
feat.ct <- perm_features[grepl("CT\\.", perm_features)]

perms.gm <- freedman_lane_perm(df_chop, feat.gm, n_perm = n_perm, n_core = n_cores, suf = "Transformed.normalised", covs = cov.gm)
perms.sa <- freedman_lane_perm(df_chop, feat.sa, n_perm = n_perm, n_core = n_cores, suf = "Transformed.normalised", covs = cov.sa)
perms.ct <- freedman_lane_perm(df_chop, feat.ct, n_perm = n_perm, n_core = n_cores, suf = "Transformed.normalised", covs = cov.ct)

perms <- bind_rows(perms.gm, perms.sa, perms.ct)

true_brain_map <- res_chop_vol %>% select(label, Cohens_d)
comp_brain_map <- res_enigma_vol %>% select(label, Cohens_d)

res <- freedman_lane_corr_test(true_brain_map, perms, comp_brain_map)
```

### Generate Plots

```{r}
#| label: define-plot-params-idp
#| output: false
fill_col <- "Cohens_d"
ci_lo_col <- "Cohens_d_ci_lo"
ci_hi_col <- "Cohens_d_ci_hi"
sig_col  <- "sig"
colorbar_title <- "Cohen's d"
max_lim <- max(abs(res_chop_vol$Cohens_d), abs(res_enigma_vol$Cohens_d))
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)
```

```{r}
#| output: false
#| label: build-brain-maps-idp

p_chop_vol <- assemble_main_figure_layout(res_chop_vol, df_chop, fill_col, color_scheme, plot_lims,
                 plot_type = "volume", colorbar_title = colorbar_title,sig_col = sig_col,
                 bar_global = T)

p_enigma_vol <- assemble_main_figure_layout(res_enigma_vol, df_enigma, fill_col, color_scheme, 
                              plot_lims,plot_type = "volume", 
                              colorbar_title = colorbar_title,sig_col = sig_col,
                 bar_global = T)
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: chop-effect-idp
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the CHOP cohort."
p_chop_vol
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: enigma-effect-idp
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the ENIGMA Cohort."
p_enigma_vol
```

#### Combine Plots

```{r, fig.width = 6.69, fig.height = 11}
#| label: combine-plots-idp
#| fig-cap: "Combined CHOP and ENIGMA case-control effect sizes (Cohen's d) using non-benchmarked volume, surface area and cortical thickness measures, with between-cohort permutation-test correlations."
t1 <- ggdraw() + 
  draw_label("CHOP", x = 0.01, hjust = 0, fontface = "bold")
t2 <- ggdraw() + 
  draw_label("ENGIMA", x = 0.01, hjust = 0, fontface = "bold")

gm_text <- glue('Gray Matter Volume: r = {round(res[res$measure == "GM","corr"][[1]],2)}, p = {round(res[res$measure == "GM","pval"][[1]],4)}')
sa_text <- glue('Surface Area: r = {round(res[res$measure == "SA","corr"][[1]],2)}, p = {round(res[res$measure == "SA","pval"][[1]],4)}')
ct_text <- glue('Cortical Thickness: r = {round(res[res$measure == "CT","corr"][[1]],2)}, p = {round(res[res$measure == "CT","pval"][[1]],4)}')

caption <- ggdraw() + 
              draw_label(glue('{gm_text} | {sa_text} | {ct_text}'), size = base_font) + figure_theme

p <- ggarrange(t1, p_chop_vol, t2, p_enigma_vol, nrow = 4, heights = c(0.08,1,0.08,1))
ggarrange(p, caption, heights = c(11.6,0.5), ncol = 1)
dir.create("../figures_final/")
ggsave("../figures_final/png/figure_s19_sensitivity_nonbenchmarked.png", bg = 'white')
ggsave("../figures_final/pdf/figure_s19_sensitivity_nonbenchmarked.pdf", bg = 'white')
```

### Describe Changes

```{r}
#| label: describe-changes-idp
# --- swap this to summarize a different sensitivity result -------------------
res_test <- res_chop_vol    # tibble to summarize
res_ref  <- res_chop_z       # reference (primary) result to compare against
# -----------------------------------------------------------------------------

# precompute the three row-wise comparisons (res_test and res_ref must be row-aligned)
sign_consistent <- sign(res_test$Cohens_d) == sign(res_ref$Cohens_d)
test_sig        <- res_test$sig == "p.fdr < 0.05"
sig_matches     <- res_test$sig == res_ref$sig

res_summary <- tibble::tibble(
  category = c(
    "Significant and directionally consistent",
    "Significant but directionally different",
    "No longer significant, but directionally consistent",
    "No longer significant and directionally different",
    "Newly significant, but directionally consistent",
    "Newly significant and directionally different",
    "Never significant and not directionally consistent",
    "Never significant, but directionally consistent"
  ),
  count = c(
    sum(sign_consistent  & test_sig  & sig_matches,  na.rm = TRUE),
    sum(!sign_consistent & test_sig  & sig_matches,  na.rm = TRUE),
    sum(sign_consistent  & !test_sig & !sig_matches, na.rm = TRUE),
    sum(!sign_consistent & !test_sig & !sig_matches, na.rm = TRUE),
    sum(sign_consistent  & test_sig  & !sig_matches, na.rm = TRUE),
    sum(!sign_consistent & test_sig  & !sig_matches, na.rm = TRUE),
    sum(!sign_consistent & !test_sig & sig_matches,  na.rm = TRUE),
    sum(sign_consistent  & !test_sig & sig_matches,  na.rm = TRUE)
  )
)

res_summary
```

## Sensitivity Analysis: Including Scans with Pathology (Figure S20)

```{r}
#| label: build-pathology-dataset
#| output: false
df_tmp <- abnormal_pathology %>% filter(!(participant %in% df_chop$participant))
df_chop_path <- bind_rows(df_chop, df_tmp)
```

### CHOP

```{r}
#| output: false
#| label: fit-models-chop-pathology
cov_str <- " + age_days + sex + euler"
rand_str <- ""

mpr_only_cols <- features[grepl("CT",features)]

res <- fit_group_comparison_models(df_chop_path, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)

res_chop_path <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-chop-pathology
print_top_significant_results(res_chop_path)
summarize_results_by_cortex(res_chop_path, caption = "Case-Control Results Grouped by Cortex in CHOP")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-chop-pathology
#| output: false
table_fmt <- fmt_table(res_chop_path)
write_csv(table_fmt, "../tables/formatted/table_s21.csv")
```

### Permutation Test

#### Generate CHOP Permutations

```{r}
#| label: run-permutation-test-pathology
#| output: false
perms <- freedman_lane_perm(df_chop_path, perm_features, n_perm = n_perm, n_core = n_cores, suf = "Transformed.q.zscore")

true_brain_map <- res_chop_path %>% select(label, Cohens_d)
comp_brain_map <- res_enigma_z %>% select(label, Cohens_d)

res <- freedman_lane_corr_test(true_brain_map, perms, comp_brain_map)
```

### Generate Plots

```{r}
#| label: define-plot-params-pathology
#| output: false
fill_col <- "Cohens_d"
ci_lo_col <- "Cohens_d_ci_lo"
ci_hi_col <- "Cohens_d_ci_hi"
sig_col  <- "sig"
colorbar_title <- "Cohen's d"
max_lim <- max(abs(res_chop_path$Cohens_d), abs(res_enigma_z$Cohens_d))
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)
```

```{r}
#| output: false
#| label: build-brain-map-pathology

p_chop_path <- assemble_main_figure_layout(res_chop_path, df_chop_path, fill_col, 
                             color_scheme, plot_lims,plot_type = "zscore", 
                             colorbar_title = colorbar_title,sig_col = sig_col)
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: chop-effect-pathology
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the CHOP cohort (including those with pathology)."
p_chop_path
```

#### Combine Plots

```{r, fig.width = 6.69, fig.height = 5}
#| label: combine-plots-pathology
#| echo: false
#| fig-cap: "Case-control effect sizes (Cohen's d) in the CHOP cohort including participants with abnormal pathology."
p_chop_path
ggsave("../figures_final/png/figure_s20_sensitivity_pathology.png", bg = 'white')
ggsave("../figures_final/pdf/figure_s20_sensitivity_pathology.pdf", bg = 'white')
```

### Describe Changes

```{r}
#| label: describe-changes-pathology
# --- swap this to summarize a different sensitivity result -------------------
res_test <- res_chop_path    # tibble to summarize
res_ref  <- res_chop_z       # reference (primary) result to compare against
# -----------------------------------------------------------------------------

summarize_model_changes(res_test, res_ref)
```

## 

## Sensitivity Analysis: Scanner Covariate (Figure S21)

*Sensitivity analysis: adding scanner/site as a fixed-effect covariate.*

### CHOP

```{r}
#| output: false
#| label: fit-models-chop-site
cov_str <- " + age_days + sex + euler + site"
rand_str <- ""

mpr_only_cols <- features[grepl("CT",features)]

res <- fit_group_comparison_models(df_chop, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)

res_chop_fixed <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-chop-site
print_top_significant_results(res_chop_fixed)
summarize_results_by_cortex(res_chop_fixed, caption = "Case-Control Results Grouped by Cortex in CHOP")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-chop-site
#| output: false
table_fmt <- fmt_table(res_chop_fixed)
write_csv(table_fmt, "../tables/formatted/table_s22.csv")
```

### ENIGMA

```{r}
#| output: false
#| label: fit-models-enigma-site
cov_str <- " + age_days + sex + euler + site"
rand_str <- ""

mpr_only_cols <- c()

res <- fit_group_comparison_models(df_enigma, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)

res_enigma_fixed <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-enigma-site
print_top_significant_results(res_enigma_fixed)
summarize_results_by_cortex(res_enigma_fixed, caption = "Case-Control Results Grouped by Cortex in ENIGMA")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-enigma-site
#| output: false
table_fmt <- fmt_table(res_enigma_fixed)
write_csv(table_fmt, "../tables/formatted/table_s23.csv")
```

### Permutation Test

#### Generate CHOP Permutations

```{r}
#| label: run-permutation-test-site
#| output: false
# ADDITIONAL COVS
perms <- freedman_lane_perm(df_chop, perm_features, n_perm = n_perm, 
                            n_core = n_cores, suf = "Transformed.q.zscore",
                            covs = "age_days + sex + euler + site")

true_brain_map <- res_chop_fixed %>% select(label, Cohens_d)
comp_brain_map <- res_enigma_fixed %>% select(label, Cohens_d)

res <- freedman_lane_corr_test(true_brain_map, perms, comp_brain_map)
```

### Generate Plots

```{r}
#| label: define-plot-params-site
#| output: false
fill_col <- "Cohens_d"
ci_lo_col <- "Cohens_d_ci_lo"
ci_hi_col <- "Cohens_d_ci_hi"
sig_col  <- "sig"
colorbar_title <- "Cohen's d"
max_lim <- max(abs(res_chop_fixed$Cohens_d), abs(res_enigma_fixed$Cohens_d))
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)
```

```{r}
#| output: false
#| label: build-brain-maps-site

p_chop_fixed <- assemble_main_figure_layout(res_chop_fixed, df_chop, fill_col, color_scheme, plot_lims,
                 plot_type = "zscore", colorbar_title = colorbar_title,sig_col = sig_col)

p_enigma_fixed <- assemble_main_figure_layout(res_enigma_fixed, df_enigma, fill_col, color_scheme,
                               plot_lims,plot_type = "zscore", 
                               colorbar_title = colorbar_title, sig_col = sig_col)
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: chop-effect-fixed
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the CHOP cohort (mixed effects model with random intercept for site)."
p_chop_fixed
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: enigma-effect-fixed
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the ENIGMA Cohort (mixed effects model with random intercept for site)."
p_enigma_fixed
```

#### Combine Plots

```{r, fig.width = 6.69, fig.height = 11}
#| label: combine-plots-site
#| fig-cap: "Combined CHOP and ENIGMA case-control effect sizes (Cohen's d) covarying for scanner site, with between-cohort permutation-test correlations."
t1 <- ggdraw() + 
  draw_label("CHOP", x = 0.01, hjust = 0, fontface = "bold")
t2 <- ggdraw() + 
  draw_label("ENGIMA", x = 0.01, hjust = 0, fontface = "bold")

gm_text <- glue('Gray Matter Volume: r = {round(res[res$measure == "GM","corr"][[1]],2)}, p = {round(res[res$measure == "GM","pval"][[1]],4)}')
sa_text <- glue('Surface Area: r = {round(res[res$measure == "SA","corr"][[1]],2)}, p = {round(res[res$measure == "SA","pval"][[1]],4)}')
ct_text <- glue('Cortical Thickness: r = {round(res[res$measure == "CT","corr"][[1]],2)}, p = {round(res[res$measure == "CT","pval"][[1]],4)}')

caption <- ggdraw() + 
              draw_label(glue('{gm_text} | {sa_text} | {ct_text}'), size = base_font) + figure_theme
              

p <- ggarrange(t1, p_chop_fixed, t2, p_enigma_fixed, nrow = 4, heights = c(0.08,1,0.08,1))
ggarrange(p, caption, heights = c(11.6,0.5), ncol = 1)
dir.create("../figures_final/")
ggsave("../figures_final/png/figure_s21_sensitivity_scanner.png", bg = 'white')
ggsave("../figures_final/pdf/figure_s21_sensitivity_scanner.pdf", bg = 'white')
```

### Describe Changes

```{r}
#| label: describe-changes-site
# --- swap this to summarize a different sensitivity result -------------------
res_test <- res_chop_fixed    # tibble to summarize
res_ref  <- res_chop_z       # reference (primary) result to compare against
# -----------------------------------------------------------------------------

summarize_model_changes(res_test, res_ref)
```

## Sensitivity Analysis: Participants Aged 6 Years and Older (Figure S22)

```{r}
#| label: build-age6plus-dataset
#| output: false
df_chop_older <- df_chop %>% filter(age_days >= (365.25*6+280))
```

### CHOP

```{r}
#| output: false
#| label: fit-models-chop-age6plus
cov_str <- " + age_days + sex + euler"
rand_str <- ""

mpr_only_cols <- features[grepl("CT",features)]
res <- fit_group_comparison_models(df_chop_older, cov_str, features, rand_str = rand_str,
                 predictor = "dx", group_control = "CN",
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)

res_chop_older <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-chop-age6plus
print_top_significant_results(res_chop_older)
summarize_results_by_cortex(res_chop_older, caption = "Case-Control Results Grouped by Cortex in CHOP")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-chop-age6plus
#| output: false
table_fmt <- fmt_table(res_chop_older)
write_csv(table_fmt, "../tables/formatted/table_s24.csv")
```

### Permutation Test

#### Generate CHOP Permutations

```{r}
#| label: run-permutation-test-age6plus
#| output: false
perms <- freedman_lane_perm(df_chop_older, perm_features, n_perm = n_perm, n_core = n_cores, suf = "Transformed.q.zscore")

true_brain_map <- res_chop_older %>% select(label, Cohens_d)
comp_brain_map <- res_enigma %>% select(label, Cohens_d)

res <- freedman_lane_corr_test(true_brain_map, perms, comp_brain_map)
```

### Generate Plots

```{r}
#| label: define-plot-params-age6plus
#| output: false
fill_col <- "Cohens_d"
ci_lo_col <- "Cohens_d_ci_lo"
ci_hi_col <- "Cohens_d_ci_hi"
sig_col  <- "sig"
colorbar_title <- "Cohen's d"
max_lim <- max(abs(res_chop_older$Cohens_d), abs(res_enigma_z$Cohens_d))
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)
```

```{r, fig.width = 6.5, fig.height = 6.5}
#| output: false
#| label: build-brain-map-age6plus

p_chop_older <- assemble_main_figure_layout(res_chop_older, df_chop_older, fill_col,
                             color_scheme, plot_lims,plot_type = "zscore",
                             colorbar_title = colorbar_title,sig_col = sig_col)
```

```{r, fig.width = 8.5, fig.height = 6.5}
#| echo: false
#| label: chop-effect-older
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the CHOP cohort (ages 6+)."
p_chop_older
```

#### Combine Plots

```{r, fig.width = 6.69, fig.height = 5}
#| label: combine-plots-age6plus
#| echo: false
#| fig-cap: "Case-control effect sizes (Cohen's d) in the CHOP cohort restricted to participants aged 6 years and older."
p_chop_older
ggsave("../figures_final/png/figure_s22_sensitivity_age6plus.png", bg = 'white')
ggsave("../figures_final/pdf/figure_s22_sensitivity_age6plus.pdf", bg = 'white')
```

### Describe Changes

```{r}
#| label: describe-changes-age6plus
# --- swap this to summarize a different sensitivity result -------------------
res_test <- res_chop_older    # tibble to summarize
res_ref  <- res_chop_z       # reference (primary) result to compare against
# -----------------------------------------------------------------------------

summarize_model_changes(res_test, res_ref)
```

## Sensitivity Analysis: Self-Described Race Covariate (Figure S23)

*Sensitivity analysis: covarying for self-described race.*

### CHOP

```{r}
#| output: false
#| label: fit-models-chop-race
cov_str <- " + age_days + sex + euler + race"
rand_str <- ""

mpr_only_cols <- features[grepl("CT",features)]

res <- fit_group_comparison_models(df_chop, cov_str, features, rand_str = rand_str, 
                 predictor = "dx", group_control = "CN", 
                 group_case = "22q11DS", model_type = "zscore",
                 mpr_only_cols = mpr_only_cols)

res_chop_race <- summarize_model_results(res,"predictorTRUE")
```

```{r}
#| echo: false
#| label: summarize-results-chop-race
print_top_significant_results(res_chop_race)
summarize_results_by_cortex(res_chop_race, caption = "Case-Control Results Grouped by Cortex in CHOP")
```

#### Generate Formatted Report Table

```{r}
#| label: save-table-chop-race
#| output: false
table_fmt <- fmt_table(res_chop_race)
write_csv(table_fmt, "../tables/formatted/table_s25.csv")
```

### Generate Plots

```{r}
#| label: define-plot-params-race
#| output: false
fill_col <- "Cohens_d"
ci_lo_col <- "Cohens_d_ci_lo"
ci_hi_col <- "Cohens_d_ci_hi"
sig_col  <- "sig"
colorbar_title <- "Cohen's d"
max_lim <- max(abs(res_chop_race$Cohens_d), abs(res_chop_race$Cohens_d))
plot_lims <- c(-1*ceiling(10*max_lim)/10,ceiling(10*max_lim)/10)
```

```{r, fig.width = 6.69, fig.height = 5}
#| output: false
#| label: build-brain-map-race

p_chop_race <- assemble_main_figure_layout(res_chop_race, df_chop, fill_col, color_scheme, plot_lims,
                 plot_type = "zscore", colorbar_title = colorbar_title,sig_col = sig_col)
```

```{r, fig.width = 6.69, fig.height = 5}
#| echo: false
#| label: chop-effect-race
#| fig-cap: "Differences in global and regional centile scores of structural MRI metrics between case and control groups in the CHOP cohort (mixed effects model with random intercept for site)."
p_chop_race
```

#### Combine Plots

```{r, fig.width = 6.69, fig.height = 5}
#| label: combine-plots-race
#| echo: false
#| fig-cap: "Case-control effect sizes (Cohen's d) in the CHOP cohort covarying for self-described race."
p_chop_race
ggsave("../figures_final/png/figure_s23_sensitivity_race.png", bg = 'white')
ggsave("../figures_final/pdf/figure_s23_sensitivity_race.pdf", bg = 'white')
```

### Describe Changes

```{r}
#| label: describe-changes-race
# --- swap this to summarize a different sensitivity result -------------------
res_test <- res_chop_race    # tibble to summarize
res_ref  <- res_chop_z       # reference (primary) result to compare against
# -----------------------------------------------------------------------------

summarize_model_changes(res_test, res_ref)
```

## Session Info

```{r}
#| label: session-info
sessionInfo()
```
